---
title: "simpleRegress"
author: "Elizabeth Rivera"
date: "Fall- NOV 16, 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

## now predict the length of a bear who has a neck of 17, 22 and 26


```{r}
new <- data.frame(NECK=c(17,22,26))
predict(simp,new)
predict(simp,new,se.fit=TRUE)
```
## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).

```{r}
simp2 <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp2)
summary.lm(simp2)
```

The adjusted R-squared is 0.5321 which is smaller than the model using the "neck."

```{r}
simp3 <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(simp3)
summary.lm(simp3)
```
The adjusted R-squared is 0.7423 which is smaller than the model using the "neck."

```{r}
simp4 <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(simp4)
summary.lm(simp4)
```
The adjusted R-squared "CHEST" variable is 0.7869 which is larger than the model using the "neck." As of now, this is the largest "chest" variable larger than the model using the "neck."


```{r}
simp5 <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(simp5)
summary.lm(simp5)
```
The adjusted R-squared "HEADLEN" variable is 0.8434 which is larger than the model using the "neck." As of now, this is the largest "HEADLEN" variable larger than the model using the "neck."


```{r}
simp6 <- lm(LENGTH~as.numeric(SEX),data=bears)
plot(LENGTH~as.numeric(SEX),data=bears)
abline(simp6)
summary.lm(simp6)
```
The adjusted R-squared "SEX" variable is -0.008373 which is too small to consider. 


```{r}
simp7 <- lm(LENGTH~as.numeric(MONTH),data=bears)
plot(LENGTH~as.numeric(MONTH),data=bears)
abline(simp7)
summary.lm(simp7)
```
The adjusted R-squared "MONTH" variable is -0.01153 which is too small to consider as well.

```{r}
simp8 <- lm(LENGTH~as.numeric(AGE),data=bears)
plot(LENGTH~as.numeric(AGE),data=bears)
abline(simp8)
summary.lm(simp8)
```
The adjusted R-squared "AGE" variable is 0.5073 which is larger than the model using the "neck." Although it is larger than the model, there already exists a value that is larger than 0.5073.

## Conclusion    

Based on Dr. K's Model LENGTH ~ NECK the R- squared variable is 0.7476. After several attempts using the simple regression model the largest R-squared variable "HEADLEN" which is 0.8434 which larger than Dr. K's example of bear length.   


